---
title: 'Code for QSS tidyverse Chapter 2: Causality'
author: "Kosuke Imai and Nora Webb Williams"
date: "First Printing"
output:
  pdf_document: default
always_allow_html: true
---

# Causality

```{r setup2, include = FALSE, purl=FALSE, eval = T}
library(knitr)
library(kableExtra)
library(tidyverse)
opts_chunk$set(tidy = FALSE,
        fig.align = "center",
        background = "#EEEEEE",
        fig.width = 7,
        fig.height = 7,
        out.width = ".7\\linewidth",
        out.height = ".7\\linewidth",
        cache = TRUE)
options(width = 60)
set.seed(12345)
```

## Racial Discrimination in the Labor Market 


```{r, message=FALSE}
## Load required packages
library(tidyverse)
library(qss)

## Load in the data from QSS package
data(resume, package = "qss")
```
```{r}
## Or read in a saved CSV
resume <- read_csv("resume.csv")
```

```{r, message=FALSE}
resume <- read_csv("resume.csv",
                   col_types = cols(
                     firstname = col_character(),
                     sex = col_character(),
                     race = col_character(),
                     call = col_number()))
```


```{r}
dim(resume)
```

```{r}
head(resume)
tail(resume)
glimpse(resume)
```


```{r}
summary(resume)
```


```{r, message=FALSE}
race.call.summary <- resume %>%
  group_by(race, call) %>% # create for each race and callback status
  count() 

race.call.summary
```

```{r, message=FALSE}
race.call.tab <- race.call.summary %>%
  pivot_wider(names_from = call, # reshape the data
              values_from = n)

race.call.tab
```

```{r}
race.call.tab.names <- race.call.tab %>% 
    rename(no_callback = "0",
         callback = "1")

race.call.tab.names
```

```{r}
race.call.tab.names <- race.call.tab.names %>% 
  mutate(total_resumes = no_callback + callback,
         callback_prop = callback / total_resumes)

race.call.tab.names
```

```{r}
overall_callback <- resume %>% 
  summarize(total_callback_rate = sum(call) / n())

overall_callback
```

```{r}
overall_callback <- resume %>% 
  summarize(total_callback_rate = mean(call))

overall_callback
```


```{r, message=FALSE}
callback_by_race <- resume %>% 
  group_by(race) %>% 
  summarize(callback_rate = mean(call))

callback_by_race
```


## Subsetting Data in R

### Logical Values and Operators

```{r}
class(TRUE)
```

```{r}
as.integer(TRUE)
as.integer(FALSE)
```


```{r}
x <- c(TRUE, FALSE, TRUE) # a vector with logical values
mean(x) # proportion of TRUEs
sum(x) # number of TRUEs
```


```{r}
FALSE & TRUE
TRUE & TRUE
```


```{r}
TRUE | FALSE
FALSE | FALSE
```

```{r}
TRUE & FALSE & TRUE
```


```{r}
(TRUE | FALSE) & FALSE # the parentheses evaluate to TRUE
TRUE | (FALSE & FALSE) # the parentheses evaluate to FALSE
```

```{r}
TF1 <- c(TRUE, FALSE, FALSE)
TF2 <- c(TRUE, FALSE, TRUE)
TF1 | TF2
TF1 & TF2
```

### Relational Operators

```{r}
4 > 3
"Hello" == "hello"  # R is case-sensitive
"Hello" != "hello"
```

```{r}
x <- c(3, 2, 1, -2, -1)
x >= 2
x != 1
```

```{r}
## logical conjunction of two vectors with logical values
(x > 0) & (x <= 2)
## logical disjunction of two vectors with logical values
(x > 2) | (x <= -1)
```

```{r}
x.int <- (x > 0) & (x <= 2) # logical vector
x.int
mean(x.int) # proportion of TRUEs
sum(x.int)  # number of TRUEs
```

### Subsetting


```{r}
## callback rate for black-sounding names
resume %>% 
  filter(race == "black") %>% # only keep observations where "race" is black
  summarize(mean(call)) # take the average of call
```
```{r}
## Subset the data with black names
resumeB <- filter(resume, race == "black")
## Calculate the mean callback rate
summarize(resumeB, mean(call))
## with $ operator to run mean() on the call column
mean(resumeB$call)
```

```{r}
resumeBf <- filter(resume, race == "black" & sex == "female")
```



```{r}
## callback rate for black female names
Bf_callback <- filter(resume, race == "black" & sex == "female") %>%
  summarize(callback_rate = mean(call)) %>% 
  pull()

## print the value to the console
print(Bf_callback)

## callback rate for black male names
Bm_callback <- filter(resume, race == "black" & sex == "male") %>%
  summarize(callback_rate = mean(call)) %>% 
  pull()

## print the value to the console
print(Bm_callback)

## callback rate for white female names
Wf_callback <- filter(resume, race == "white" & sex == "female") %>%
  summarize(callback_rate = mean(call)) %>% 
  pull()

## print the value to the console
print(Wf_callback)

## callback rate for white male names
Wm_callback <- filter(resume, race == "white" & sex == "male") %>%
  summarize(callback_rate = mean(call)) %>% 
  pull()

## print the value to the console
print(Wm_callback)

## difference between white women and black women
Wf_callback - Bf_callback

## difference between white men and black men
Wm_callback - Bm_callback
```

 

```{r, message=FALSE}
racial_gaps_by_sex <- resume %>% 
  group_by(race, sex) %>% # using two variables to group the data
  summarize(callback = mean(call)) %>% # the callback rate for each group
  pivot_wider(names_from = race, # reshaping the data
              values_from = callback) %>% 
  mutate(race_gap = white - black)

print(racial_gaps_by_sex)
```


```{r, message=FALSE}
## what happens in this portion of the code?
resume %>% 
  group_by(race, sex) %>% 
  summarize(callback = mean(call))

## What happens after we add the pivot_wider()? 
resume %>% 
  group_by(race, sex) %>% 
  summarize(callback = mean(call)) %>% 
  pivot_wider(names_from = race,
              values_from = callback)

## And so on
```

### Simple Conditional Statements

```{r}
resume <- resume %>% 
  ## create a new variable that is 1 if the resume is black and female, 
  ## 0 otherwise
  mutate(BlackFemale = if_else(race == "black" & 
                               sex == "female", 1, 0))
```

```{r}
# Rows in the resumeBf data
nrow(resumeBf)
# Is that equal to sum of BlackFemale?
nrow(resumeBf) == resume %>% summarize(bf = sum(BlackFemale))
```


### Factor Variables


```{r}
resume_fact <- resume %>% 
  mutate(type = if_else(race == "black" & sex == "female", "BlackFemale", ""),
         type = if_else(race == "black" & sex == "male", "BlackMale", type),
         type = if_else(race == "white" & sex == "female", "WhiteFemale", type),
         type = if_else(race == "white" & sex == "male", "WhiteMemale", type))

head(resume_fact)
```


```{r}
resume <- resume %>% 
  ## add a categorical variable for race/gender type
  mutate(type = case_when(race == "black" & sex == "female" ~ "BlackFemale",
                          race == "white" & sex == "female" ~ "WhiteFemale",
                          race == "black" & sex == "male" ~ "BlackMale",
                          race == "white" & sex == "male" ~ "WhiteMale",
                          TRUE ~ "other"
))

head(resume)

## Did any observations receive the "other" value for type?
filter(resume, type == "other")
```

```{r}
## check object class
class(resume$type)

## coerce the character variable into a factor variable
resume <- resume %>% 
  mutate(type = as.factor(type))

## look at the levels of the factor
levels(resume$type)
```

```{r, message=FALSE}
firstname_callback <- resume %>% 
  group_by(firstname) %>% 
  select(firstname, call) %>% 
  summarize(callback = mean(call))

head(firstname_callback)
```

## Causal Effects and the Counterfactual


```{r}
slice(resume, 1)
```


## Randomized Controlled Trials 

### The Role of Randomization

### Social Pressure and Voter Turnout 

```{r}
## Load in the data from QSS package
data(resume, package = "qss")

## Or from a csv
social <- read_csv("social.csv", 
                   col_types = cols(sex = col_character(),
                                     yearofbirth = col_double(),
                                     primary2004 = col_double(),
                                     messages = col_character(),
                                     primary2006 = col_double(),
                                     hhsize = col_double())) 
summary(social) # summarize the data
```


```{r, message=FALSE}
## Average turnout by treatment message
turnout_by_message <- social %>%
  group_by(messages) %>%
  summarize(turnout = mean(primary2006))

turnout_by_message

## Differences between treatment(s) and control means
turnout_diffs <- turnout_by_message %>% 
  pivot_wider(names_from = messages,
              values_from = turnout) %>% 
  mutate(diff_Civic_Duty = `Civic Duty` - Control,
         diff_Hawthorne = Hawthorne - Control,
         diff_Neighbors = Neighbors - Control) %>% 
  select(diff_Civic_Duty, diff_Hawthorne, diff_Neighbors)

turnout_diffs
```

```{r, message=FALSE}
social %>% 
  mutate(age = 2006 - yearofbirth) %>% 
  group_by(messages) %>% 
  summarize(age_avg = mean(age),
            primary2004_avg = mean(primary2004),
            hhsize_avg = mean(hhsize))
```

## Observational Studies

### Minimum Wage and Unemployment

```{r, message=FALSE}
minwage <- read_csv("minwage.csv") # load the data
## or 
data(minwage, package = "qss")
dim(minwage) # dimension of data
glimpse(minwage)
summary(minwage) # summary of data
```


```{r, message=FALSE}
## Add a 'state' variable
minwage <- minwage %>% 
  mutate(state = if_else(location == "PA", "PA", "NJ"))

## Create the 'new_wage' object
new_wage <- 5.05

## Calculate the proportions above and below the new wage by state
state_props <- minwage %>% 
  mutate(above_min_before = if_else(wageBefore >= new_wage, 1, 0),
         above_min_after = if_else(wageAfter >= new_wage, 1, 0)) %>% 
  group_by(state) %>% 
  summarize(prop_before = mean(above_min_before),
            prop_after = mean(above_min_after))

state_props
```


```{r, message=FALSE}
## First create new variables to calculate the 
## proportion of full-time employees
minwage <- minwage %>%
  mutate(totalAfter = fullAfter + partAfter,
        fullPropAfter = fullAfter / totalAfter)

## Then calculate the average proportion of full-time workers by state 
full_prop_by_state <- minwage %>%
  group_by(state) %>%
  summarize(fullPropAfter = mean(fullPropAfter))

## To calculate the difference between states, we use pivot_wider()
## and mutate()
pivot_wider(full_prop_by_state, 
            names_from = state, values_from = fullPropAfter) %>%
  mutate(diff = NJ - PA)
```

### Confounding Bias

```{r, message = FALSE}
chains_by_state <- minwage %>%
  group_by(state) %>%
  count(chain) %>%
  mutate(prop = n / sum(n)) %>% 
  pivot_wider(-n, # this drops the 'n' variable prior to pivoting 
              names_from = state,
              values_from = prop)

chains_by_state
```


```{r message=FALSE, warning=FALSE}
full_prop_by_state_chain <- minwage %>%
  group_by(state, chain) %>%
  summarize(fullPropAfter = mean(fullPropAfter)) %>% 
  pivot_wider(names_from = state,
              values_from = fullPropAfter) %>% 
  mutate(diff = NJ - PA)

full_prop_by_state_chain
```

```{r, message = FALSE}
prop_by_state_chain_location_subset <- minwage %>%
  filter(!location %in% c("shoreNJ", "centralNJ")) %>% 
  group_by(state, chain) %>%
  summarize(fullPropAfter = mean(fullPropAfter)) %>% 
  pivot_wider(names_from = state,
              values_from = fullPropAfter) %>% 
  mutate(diff = NJ - PA)

prop_by_state_chain_location_subset
```

### Before-and-After and Difference-in-Differences Designs

```{r}
## First, create a variable for the full-time 
## proportion prior to the change
minwage <- minwage %>%
  mutate(totalBefore = fullBefore + partBefore,
         fullPropBefore = fullBefore / totalBefore)

## Then look at the differences in average proportion of full-time
## before and after (in NJ only)
minwage %>%
  filter(state == "NJ") %>%
  summarize(diff = mean(fullPropAfter) - mean(fullPropBefore))
```

```{r, message=FALSE}
## DiD estimate
minwage %>%
  group_by(state) %>%
  ## difference before and after
  summarize(diff = mean(fullPropAfter) - mean(fullPropBefore)) %>%
  pivot_wider(names_from = state, values_from = diff) %>%
  ## difference in difference between states
  mutate(diff_in_diff = NJ - PA)
```


## Descriptive Statistics for a Single Variable

### Quantiles 

```{r, message=FALSE}
## median difference-in-differences
minwage %>%
  group_by(state) %>%
  summarize(diff = median(fullPropAfter) - median(fullPropBefore)) %>%
  pivot_wider(names_from = state, values_from = diff) %>%
  mutate(diff_in_diff = NJ - PA)
```

```{r}
## summary() shows quartiles for the two wages variables
## as well as minimum, maximum, and mean
minwage %>%
  filter(state == "NJ") %>% # just look at NJ
  select(wageBefore, wageAfter) %>%
  summary()

## The interquartile range
minwage %>%
  filter(state == "NJ") %>%
  select(wageBefore, wageAfter) %>%
  summarize(wageBeforeIQR = IQR(wageBefore),
            wageAfterIQR = IQR(wageAfter))
```


```{r}
## Create an object for the quantiles we want (deciles)
decile_probs <- seq(from = 0, to = 1, by = 0.1)
## Save deciles as characters
decile_names <- as.character(decile_probs)

## Generate the deciles for wage before and after
minwage %>% 
  filter(state == "NJ") %>%
  select(wageBefore, wageAfter) %>% 
  summarize(wageBeforeDecile = quantile(wageBefore, probs = decile_probs),
            wageAfterDecile = quantile(wageAfter, probs = decile_probs),
            decile = decile_names) 
```

### Standard Deviation 


```{r}
## Calculate the RMS of the change in full-time 
## employment proportion in NJ
minwageNJ %>% 
  mutate(fullPropChange = fullPropAfter - fullPropBefore,
         sqfullPropChange = fullPropChange^2) %>% 
  summarize(rms = sqrt(mean(sqfullPropChange)))

## Compare to the mean
minwageNJ %>% 
  mutate(fullPropChange = fullPropAfter - fullPropBefore) %>% 
  summarize(mean = mean(fullPropChange))
```


```{r, include=FALSE}
## Original, before corrections
minwage %>%
  group_by(state) %>%
  summarize_at(vars(wageAfter, wageBefore), 
               .funs = list(sd, var))

minwage %>%
  group_by(state) %>%
  summarize_at(vars(wageAfter, wageBefore), 
               .funs = list(stdv = sd, 
                            variance = var))
```

```{r}
minwage %>%
  group_by(state) %>%
  summarize_at(vars(fullPropBefore, fullPropAfter), 
               .funs = list(sd, var))

minwage %>%
  group_by(state) %>%
  summarize_at(vars(fullPropBefore, fullPropAfter), 
               .funs = list(stdv = sd, 
                            variance = var))
```

