---
title: 'Errata for *Quantitative Social Science: An Introduction in tidyverse* (Princeton University Press)'
author: "Kosuke Imai and Nora Webb Williams"
date: "March 4, 2023"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Corrections applicable to first printing

## Chapter 3

### Section 3.3.3

* page 103. In the box plot "Distribution of age," the blue curly bracket indicating the IQR should extend from the upper quartile to the lower quartile instead of from the median to the lower quartile. In addition, the blue curly bracket indicating 1.5 $\times$ IQR should extend from the upper quartile instead of from the median. 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
## remake the plot with correct curly brackets
library(tidyverse)
library(qss)

## Load in the data from QSS package
data(afghan, package = "qss")

# with assist from https://stackoverflow.com/questions/35633239/add-curly-braces-to-ggplot2-and-then-use-ggsave
library(pBrackets)
bracketsGrob <- function(...){
l <- list(...)
e <- new.env()
e$l <- l
  grid:::recordGrob(  {
    do.call(grid.brackets, l)
  }, e)
}

b1 <- bracketsGrob(.75, .39, .75, .14, h=0.06, lwd=1.5, col="blue")
b2 <- bracketsGrob(.75, .77, .75, .39, h=0.06, lwd=1.5, col="blue")

ggplot(afghan, aes(y = age)) +
  geom_boxplot() +
  labs(y = "Age", x = "", title = "Distribution of age") +
  xlim(-.75, .75) +
  ggplot2::annotate("text", x = -.56, y = 40, label = "Upper quartile", size = 14/.pt) +
  ggplot2::annotate("text", x = -.56, y = 23, label = "Lower quartile", size = 14/.pt) +
  ggplot2::annotate("text", x = -.56, y = 31, label = "Median", size = 14/.pt) +
  annotation_custom(b1) +
  annotation_custom(b2) +
  ggplot2::annotate("text", x = .48, y = 53, label = "1.5 x IQR", 
           hjust = -.15,
           color = "blue", size = 14/.pt) +
  ggplot2::annotate("text", x = .48, y = 31, label = "IQR", 
           hjust = -.25,
           color = "blue", size = 14/.pt) +
  theme_classic(base_size = 12) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(plot.title = element_text(hjust = 0.5))
```

### Section 3.6.2

* pages 120-121. When describing the Gini coefficient in both text and text box, 0 and 1 should be flipped. Text should read "In a perfectly equal society, the Gini coefficient is 0. In contrast, a society where one person possesses all the wealth has a Gini coefficient of 1." Text box should read "It ranges from 0 (everyone has the same amount of wealth) to 1 (one person possesses all the wealth)."

## Chapter 7

### Section 7.1.3

* page 372. The in-text equation, $P(Z > z_{\alpha/2})=1-P(Z\le z_{\alpha/2})=1-\alpha/2$, should be $P(Z > z_{\alpha/2})=1-P(Z\le z_{\alpha/2})=\alpha/2$.

## Acknowledgements

Thanks to Shiro Kuriwaki, Danya Lagos, and Sanghoon Kim-Leffingwell for pointing out errors. 

